---
title: "Mushroom Data Transformation"
author: "Venkatesan Parkunan"
date: "9/3/2018"
output: 
    html_document:
    theme: lumen
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This assignment for DATA 607 course within the CUNY School of Professional Studies MS Data Science program is to transform semi-unstructured data into a structured data using R methods and functions:

#Loading dplyr package

```{r include=FALSE}
library(readr)
library(dplyr)
library(plyr)
mushroom <- read_csv("https://raw.githubusercontent.com/vepark/Datasets/mushroom/mushroom.txt", col_names = FALSE)
mushroom
```

#Copying the mushroom dataframe to df dataframe 
```{r}
df <- mushroom 
```

#Selecting only specific columns to work with and writing it to df1 dataframe
``` {r}
df1 <- df[c("X1","X2","X3","X4")]
```
#Renaming the df1 dataframe columns to the right column names
```{r}
df2 <- rename(df1, c("X1"="Classes","X2"="Cap_Shape",	"X3"="Cap_Surface",	"X4"="Cap_color"))
df2  
```
#Changing the values of the dataframe for the above variables
```{r}
df2$Classes[df2$Classes == "e"] <- "Edible"
df2$Classes[df2$Classes == "p"] <- "Poisonous"

df2$Cap_Shape[df2$Cap_Shape == "b"] <- "Bell"
df2$Cap_Shape[df2$Cap_Shape == "c"] <- "Conical"
df2$Cap_Shape[df2$Cap_Shape == "x"] <- "Convex"
df2$Cap_Shape[df2$Cap_Shape == "f"] <- "Flat"
df2$Cap_Shape[df2$Cap_Shape == "k"] <- "Knobbed"
df2$Cap_Shape[df2$Cap_Shape == "s"] <- "Sunken"

df2$Cap_Surface[df2$Cap_Surface == "f"] <- "Fibrous"
df2$Cap_Surface[df2$Cap_Surface == "g"] <- "Grooves"
df2$Cap_Surface[df2$Cap_Surface == "y"] <- "Scaly"
df2$Cap_Surface[df2$Cap_Surface == "s"] <- "Smooth"

df2$Cap_color[df2$Cap_color == "n"] <- "Brown"
df2$Cap_color[df2$Cap_color == "b"] <- "Buff"
df2$Cap_color[df2$Cap_color == "c"] <- "Cinnamon"
df2$Cap_color[df2$Cap_color == "g"] <- "Gray"
df2$Cap_color[df2$Cap_color == "r"] <- "Green"
df2$Cap_color[df2$Cap_color == "p"] <- "Pink"
df2$Cap_color[df2$Cap_color == "u"] <- "Purple"
df2$Cap_color[df2$Cap_color == "e"] <- "Red"
df2$Cap_color[df2$Cap_color == "w"] <- "White"
df2$Cap_color[df2$Cap_color == "y"] <- "Yellow"

FinalMushroomData <- df2
FinalMushroomData
```